Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-08 csf-2017/paper.tex
lr

@@

The properties of the \emph{private future region} relation are:
\begin{mathpar}
  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
              (s,s') \in \phi} 
  {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr \revoked }
\end{mathpar}
These rules express that, in the private future region relation, 
we allow revocation of temporary regions. 
As explained above, this can be done using a special revoked region,
which does not desribe any memory, but we also allow  
any other region to serve as a mask, and thus any region $r$ can be a
future region of a temporary region and of a revoked region.
If a $\temp$ region is not masked by a region, then it behaves like a
$\perm$ region. The $\perma$ regions stay permanent and they are
allowed to take a transition in the private part of the transition
system.

The \emph{public future} region relations satisfies the following properties:
\begin{mathpar}
  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
    (s,s') \in \phi_\pub }
  {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
  \and
  \inferrule{ }
  { \revoked \futurewk \revoked }
\end{mathpar}
In the public future regions we do not permit revocation of temporary
regions. Instead, temporary and permanent regions are only 
allowed to take a transition in the public part of the transition system.

The $\revoked$ regions serve as masking regions, but are not as
liberal as in the private future region relation when it comes to
letting other regions replace them as masks. In the public future
region relation, we only allow $\revoked$ and $\temp$ regions to take
on the job as masking region. We have this limitation, because we
want the private future region relation to be able to reuse any region
it has revoked.
\todo[inline]{can we make this crisper ?}

@@

@@

\subsubsection{Observation relation}
\todo[inline]{write something about single orthogonal closure?}  The
\emph{observation relation}, $\observations$, is a world-indexed
predicate that captures all the configurations that produce an
admissable result. Here we take the result of a computation to be the
memory when the machine halts successfully. We expect the memory to
adhere to the protocols specified by the world, but the memory may
change during the execution, so we just want the resulting memory to
satisfy some private future world.

We also consider a failing execution admissable, because the assembly
language contains a $\fail$ instruction, and we want to be able to
reason about abritrary adverserial programs.

The observation relation is defined as follows:

\begin{align*}
  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
  \observations (W) \defeq & \lambda W \ldotp 
                             \{ \npair{(\reg,\ms)} \mid
                             \begin{aligned}[t]
                               & \forall \ms_f, \heap', i \leq n \ldotp \\
                               & \quad(\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap')  \\
                               & \qquad \Rightarrow
                               \begin{aligned}[t]
                                 & \exists W' \futurestr W \ldotp\exists \ms_r, \ms' \ldotp\\
                                 & \quad \heap' = \ms' \uplus \ms_r \uplus \ms_f \land \\ 
                                 & \quad \heapSat[\hs']{n-i}{W'} \}
                               \end{aligned}
                             \end{aligned}
\end{align*}

\subsubsection{Expression and register-file relation}
The expression relation, $\stder$, is a world-indexed predicate that
contains all the words that safely can be put in the $\pcreg$-register
of a register-file that only gives access to capabilities that are capability
safe. 
\begin{align*}
  \stder : & \Worlds \nefun \UPred{\Words} \\
  \stder \defeq & \lambda W \ldotp \{ \npair{\pc} \mid 
                  \begin{aligned}[t]
                    & \forall n' \leq n \ldotp\\
                    & \quad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
                    & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
                    & \quad \qquad \quad \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
                  \end{aligned}
\end{align*}

@@

Now define the value relation as follows:
\begin{align*}
  &\stdvr : \Worlds \monwknefun \UPred{\Words} \\
  &\stdvr\defeq \lambda \; W \ldotp 
    \begin{aligned}[t]
      & \{ \npair{i} \mid i \in \ints \} 
      \union \\
      & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
      \union \\
      & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)\} 
      \union \\
      & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}
      \union \\
      & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \}
      \union \\
      & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
      & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \} 
      \union \\
      & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)\} 
      \union \\
      & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \land\\
      & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W)  \land \\
      & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
      \union \\
      & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \land\\
      & \quad \npair{(\rwlx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
      & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
      & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
    \end{aligned}
\end{align*}

>>>>

\emph{memory}  we don't make a distinction between memory and memory
segments in the text  (to do so properly, we would need to explain why
assumptions
are reasonable, given that a real machine has finite memory, but we
need infinite domain for malloc spec, etc.)

@@

The properties of the \emph{private future region} relation are:
\begin{mathpar}
  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
              (s,s') \in \phi} 
  {  (v',s',\phi_\pub',\phi',H') \futurestr (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr (\temp,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ r \in \Regions }
  { r \futurestr \revoked }
\end{mathpar}
These rules express that, in the private future region relation, 
we allow revocation of temporary regions. 
As explained above, this can be done using a special revoked region,
which does not desribe any memory, but we also allow  
any other region to serve as a mask, and thus any region $r$ can be a
future region of a temporary region and of a revoked region.
If a temporary region is not masked by a region, then it behaves like a
permanent region. The permanent regions stay permanent and they are
allowed to take a transition in the private part of the transition
system.

The \emph{public future} region relations satisfies the following properties:
\begin{mathpar}
  \inferrule{ (v,\phi_\pub,\phi,H) = (v',\phi_\pub',\phi',H') \\
    (s,s') \in \phi_\pub }
  {  (v',s',\phi_\pub',\phi',H') \futurewk (v,s,\phi_\pub,\phi,H) }
  \and
  \inferrule{ (\temp,s,\phi_\pub,\phi,H) \in \Regions }
  { (\temp,s,\phi_\pub,\phi,H) \futurewk \revoked }
  \and
  \inferrule{ }
  { \revoked \futurewk \revoked }
\end{mathpar}
In the public future regions we do not permit revocation of temporary
regions. Instead, temporary and permanent regions are only 
allowed to take a transition in the public part of the transition system.
In the public future region relation, we only allow revoked and
temporary regions to mask a region. This means that the private future
world relations allows us to reuse a region that has been revoked
earlier.

@@

A memory satisfies a world, written $\memSat{\ms}{W}$, 
if it can be partitioned into disjoint parts such that each part is
in the interpretation of an active (permanent or temporary) region.

@@

\subsubsection{Observation relation}
\todo[inline]{write something about single orthogonal closure?}

 The \emph{observation relation}, $\observations$, defines which
configurations are capability safe. Roughly speaking,
a configuration is capability safe if, whenever it halts, then
the resulting memory satisfies a private future world. 
Notice that if a configuation fails, e.g., if an unauthorized access 
(e.g., a load of a memory location without the requisite read capability)
is attempted, then the configuation is considered capability safe, which it should
because then the capability machine has caught the attempted unauthorized access
This is similar to the logical relation given for an untyped language
in~\cite{Devriese:2016ObjCap}, but unlike logical relations for typed languages, which
typically require that computations do not fail.

\subsubsection{Expression and register-file relation}
The expression relation, $\stder$, is a world-indexed predicate that
contains all the words that safely can be put in the $\pcreg$-register
of a register-file that only gives access to capabilities that are capability
safe.

\begin{figure}[htbp]
  \centering
\begin{align*}
  \observations : &  \Worlds \nefun \UPred{\Regs \times \HeapSegments} \\
  \observations (W) \defeq & \lambda W \ldotp 
                             \{ \npair{(\reg,\ms)} \mid
                             \begin{aligned}[t]
                               & \forall \ms_f, \heap', i \leq n \ldotp \\
                               & \quad(\reg,\ms \uplus \ms_f) \step[i] (\halted,\heap')  \\
                               & \qquad \Rightarrow
                               \begin{aligned}[t]
                                 & \exists W' \futurestr W \ldotp\exists \ms_r, \ms' \ldotp\\
                                 & \quad \heap' = \ms' \uplus \ms_r \uplus \ms_f \land \\ 
                                 & \quad \heapSat[\hs']{n-i}{W'} \}
                               \end{aligned}
                             \end{aligned}
\end{align*}
\begin{align*}
  \stder : & \Worlds \nefun \UPred{\Words} \\
  \stder \defeq & \lambda W \ldotp \{ \npair{\pc} \mid 
                  \begin{aligned}[t]
                    & \forall n' \leq n \ldotp\\
                    & \quad \forall \npair[n']{\reg} \in \stdrr(W) \ldotp \\
                    & \quad \qquad  \forall \heapSat[\hs]{n'}{W} \ldotp \\
                    & \quad \qquad \quad \npair[n']{(\reg\update{\pcreg}{\pc},\hs)} \in \observations(W) \}
                  \end{aligned}
\end{align*}
\begin{align*}
  &\stdvr : \Worlds \monwknefun \UPred{\Words} \\
  &\stdvr\defeq \lambda \; W \ldotp 
    \begin{aligned}[t]
      & \{ \npair{i} \mid i \in \ints \} 
      \union \\
      & \{ \npair{\stdcap[(\noperm,\gl)] }  \} 
      \union \\
      & \{ \npair{\stdcap[(\readonly,\gl)] } \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W)\} 
      \union \\
      & \{ \npair{\stdcap[(\readwrite,\gl)] } \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \}
      \union \\
      & \{ \npair{\stdcap[(\readwritel,\gl)] } \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \}
      \union \\
      & \{ \npair{\stdcap[(\exec,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
      & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \} 
      \union \\
      & \{ \npair{\stdcap[(\entry,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend,\addr)} \in \entryCond{}(\gl)(W)\} 
      \union \\
      & \{ \npair{\stdcap[(\rwx,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\nwl,\gl)(W) \land\\
      & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W)  \land \\
      & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
      \union \\
      & \{ \npair{\stdcap[(\rwlx,\gl)]} \mid \\
      & \quad \npair{(\start,\addrend)} \in \readCond{}(\gl)(W) \land \\
      & \quad \npair{(\start,\addrend)} \in \writeCond{}(\iota^\pwl,\gl)(W) \land\\
      & \quad \npair{(\rwlx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
      & \quad \npair{(\rwx,\start,\addrend)} \in \execCond{}(\gl)(W) \land \\
      & \quad \npair{(\exec,\start,\addrend)} \in \execCond{}(\gl)(W) \}
    \end{aligned}
\end{align*}
  \caption{Logical Relation}
  \label{fig:logrel}
\end{figure}

@@

Now define the value relation as follows:
